---
title: "Lecture 2"
editor: visual
---

```{r}
library(tidyverse)

ghg <- read_csv(
  "https://bcdanl.github.io/data/GHG_emissions_by.csv"
)

positive_weak <- read_csv(
    "https://bcdanl.github.io/data/UStweets_positive_weak_2017.csv"
  )

tweets <- bind_rows(
  positive_weak = read_csv(
    "https://bcdanl.github.io/data/UStweets_positive_weak_2017.csv"
  ),
  positive_strong = read_csv(
    "https://bcdanl.github.io/data/UStweets_positive_strong_2017.csv"
  ),
  negative_weak = read_csv(
    "https://bcdanl.github.io/data/UStweets_negative_weak_2017.csv"
  ),
  negative_strong = read_csv(
    "https://bcdanl.github.io/data/UStweets_negative_strong_2017.csv"
  ),
  .id = "sample"
)
```


# Classwork 1

```{r}
library(tidyverse)

ghg_url <- "https://bcdanl.github.io/data/GHG_emissions_by.csv"

ghg <- read_csv(
  ghg_url,
  show_col_types = FALSE
)
```

One intended row represents one ________________________________ record.


## 🧪 Part 1 · Audit before answering


### 1A. Predict the output grain


```{r}
ghg |>
  summarise(
    rows = n(),
    parties = n_distinct(Party),
    first_year = min(Year),
    last_year = max(Year),
    sectors = n_distinct(Sector),
    missing_values = sum(is.na(GHG_emissions))
  )
```



### 1B. Audit the intended key


```{r}
ghg |>
  count(Party, Year, Sector) |>
  filter(n > 1)
```



### 🕳️ Part 2 · Locate missingness
```{r}
missing_by_sector <- ghg |>
  group_by(Sector) |>
  summarise(
    rows = n(),
    missing = sum(is.na(GHG_emissions)),
    missing_share = mean(is.na(GHG_emissions)),
    # .groups = "drop"
  ) |>
  arrange(desc(missing_share))

missing_by_sector
```

1. Why would `sum(GHG_emissions, na.rm = TRUE)` be different from claiming that every missing value equals zero?
2. Which additional variable would you group by to learn whether missingness is concentrated in particular years or parties?

```{r}

missing_by_sector_year <- ghg |>
  group_by(Sector, Year) |>
  summarise(
    rows = n(),
    missing = sum(is.na(GHG_emissions)),
    missing_share = mean(is.na(GHG_emissions)),
    # .groups = "drop"
  ) |>
  arrange(desc(missing_share)) |> 
  filter( missing > 0) |> 
  group_by(Sector) |> 
  summarize(mean_missing = mean(missing),
            sd_missing = sd(missing))

missing_by_sector_year
```
```{r}

missing_by_sector_party <- ghg |>
  group_by(Sector, Party) |>
  summarise(
    rows = n(),
    missing = sum(is.na(GHG_emissions)),
    missing_share = mean(is.na(GHG_emissions)),
    # .groups = "drop"
  ) |>
  arrange(desc(missing_share)) |> 
  filter( missing > 0) |> 
  group_by(Sector) |> 
  summarize(mean_missing = mean(missing),
            sd_missing = sd(missing, na.rm = T))

missing_by_sector_party

```


## 🗂️ Part 3 · Make the working table legible

```{r}
ghg_working <- ghg |>
  select(Party, Year, Sector, GHG_emissions) |>
  rename(
    party = Party,
    year = Year,
    sector = Sector,
    emissions = GHG_emissions
  ) |>
  relocate(sector, party, year)

glimpse(ghg_working)

ghg_working
```



## 🇺🇸 Part 4 · Focus the environmental question

Between 2005 and 2017, which U.S. sector values changed the most, and which changes can be expressed as percentages without using a nonpositive denominator?


```{r}
us_endpoints <- ghg_working |>
  filter(
    party == "United States of America",
    year %in% c(2005, 2017)
  ) |>
  arrange(sector, year)

us_endpoints
```

Before running it, predict:

number of rows: ________
one row now represents: ________________________________




## 🧮 Part 5 · Change the grain and calculate change

```{r}
us_change <- us_endpoints |>
  group_by(sector) |>
  summarise(
    start = first(emissions),
    end = last(emissions),
    endpoint_values = sum(!is.na(emissions)),
    .groups = "drop"
  ) |>
  mutate(
    change = end - start,
    pct_change = if_else(
      start > 0,
      100 * change / start,
      NA_real_
    )
  ) |>
  arrange(-pct_change)

us_change
```


1. What does one row in us_change represent?
2. Why did order matter before first() and last()?
3. Why is pct_change missing for LULUCF even though both endpoints exist?
4. Why should the positive change for a negative LULUCF value not be interpreted without sector documentation?


## 📊 Part 6 · Produce the analytical artifact

```{r}
final_table <- us_change |>
  transmute(
    sector,
    value_2005 = round(start),
    value_2017 = round(end),
    absolute_change = round(change),
    percent_change = round(pct_change, 1),
    audit_status = case_when(
      endpoint_values < 2 ~ "missing endpoint",
      start <= 0 ~ "nonpositive baseline",
      TRUE ~ "percent reported"
    )
  )

final_table
```


In this inventory file, the reported U.S. ________ value [rose/fell] by ________ between 2005 and 2017 (________% using a positive 2005 baseline); the file does not document ________, so the result should be interpreted as ________________________________.

